---
title: "applied_ps_2"
author: "Aditya Burra, Eva Liu & Duoduo Qi"
date: "02/05/2022"
output:
  html_document:
    number_sections: yes
  pdf_document: default
---
```{r message=FALSE, warning=FALSE}
set.seed(163264)
rm(list = ls())
library(tidyverse)
library(viridis)
library(testthat)
library(readxl)
```

<!-- .Rmd files use  markdown, a text mark up language, to provide formating.--> 
<!--Text include within these strange arrows are comments and will not show up when you knit-->

**Front matter**
This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: **H.L.**

Add your collaborators: **Duoduo Qi, Aditya Burra**

Late coins used this pset: 0. Late coins left: 4. 


# 2 Data Ingestion

## 2.1

```{r}
setwd("/Users/8q48/Documents/GitHub/applied-ps-2-aditya-joy-eva")
provider <- read_csv("unformatted_medicare_post_acute_care_hospice_by_provider_and_service_2014_12_31.csv")
problems(provider)
```

```{r}
test_that(
  "we have the right number of rows",
  expect_equal(nrow(provider), 31665)
)
```

## 2.2

```{r}
provider_hhrg <- read_excel("Provider_by_HHRG_PUF_2014.xlsx", sheet = 3)
```

```{r}
test_that(
  "we have the right number of rows",
  expect_equal(nrow(provider_hhrg), 111904)
)
```

## 2.3
```{r}
case_mix_weights <- read_csv("CY 2014 Final HH PPS Case-Mix Weights.csv")
```
```{r}
case_mix_weights <- rename(case_mix_weights,
  casemix_2014 = "2014 Final HH PPS Case-Mix Weights",
  casemix_2013 = "2013 HH PPS Case-Mix Weights"
)
```

```{r}
case_mix_weights <- case_mix_weights %>%
  select(!casemix_2013)
```

```{r}
test_that(
  "we have the right number of rows",
  expect_equal(nrow(case_mix_weights), 153)
)
```

# 3 Data Orientation and Validation

## 3.1 
* HH stands for Home Health, IRF stands for Inpatient Rehabilitation Facility, HOS stands for Hospice, LTC stands for Long Term Care, and SNF stands for Skilled Nursing Facility. 

## 3.2
* The three levels of aggregation provided are National, State, and Provider level. 

## 3.3
* According to the dataset, we see that 3,416 thousand people received home healthcare benefits in 2014.

* This website (https://www.cms.gov/Research-Statistics-Data-and-Systems/Statistics-Trends-and-Reports/CMS-Statistics-Reference-Booklet/Downloads/2015CMSStatistics.pdf) states that an approximate 3,601 thousand people got home healthcare benefits in 2014. If this is true, then the number from the dataset is inadvertantly wrong. This misalignment might be due to the units used for aggregation of beneficiaries in Bene_Dstnct_Cnt. 

```{r}
provider %>%
  filter(Srvc_Ctgry == "HH" & Prvdr_ID == "NATION") %>%
  select(Bene_Dstnct_Cnt)
```

## 3.4
* The numbers from the two sources do not align and the first source has a larger number of episodes. 
* The Methodology document states that provider_hhrg dataset does not include low utilization payment adjustment, so it is plausible that provider_hhrg has lower number of episodes.
```{r}
provider %>%
  filter(Srvc_Ctgry == "HH") %>%
  summarize(tot_epsd = sum(Tot_Epsd_Stay_Cnt))
```

```{r}
parse_hhrg <- as.data.frame(parse_number(provider_hhrg$Tot_Epsd_Stay_Cnt))
sum(parse_hhrg)
```

## 3.5
* Provider id plus Grpng is the unique identifier.
```{r}
provider_level <- provider_hhrg %>%
  filter(Smry_Ctgry == "PROVIDER")

unique <- provider_level %>%
  distinct(Prvdr_ID, Grpng)

test_that(
  "The pairs of Provider ID and group don't appear twice",
  expect_equal(nrow(unique), nrow(provider_level))
)
```

# 4 Merge Provider Costs with Provider Case-Mix

## 4.1
The str_sub command is used for extracting parts of a string.  
The required arguments are start and end arguments which give the (inclusive) position of the substring.  
The str_trim command removes whitespace from start and end of string.


## 4.2
Episodes, Therapy Visits, Clinical Levels, Functional Levels and Service levels are five types of information available in both datasets.  
(1) Episodes: It is the episode in the illness. Early episode includes 1st and 2nd episodes. Late episode includes 3rd+ episodes.  
(2) Therapy Visits: It is the number of therapy visits delivered within the 60-day episode. It includes '0-13 therapies', '14-19 therapies' and '20+ therapies'.  
(3) Clinical Levels: It means clinical severity levels. It includes level 1-3.  
(4) Functional Levels: It means functional severity levels. It includes level 1-3.  
(5) Service levels: It means service severity levels. It includes level 1-5.  


## 4.3
* I plan to use Grpng_Desc for joining.
```{r}
provider_hhrg %>%
  count(Grpng_Desc) %>%
  nrow()
```
* There are 153 distinct HHRG group using this column.

## 4.4

```{r}
new_pro_hhrg <- provider_hhrg %>%
  separate(
    Grpng_Desc,
    into = c("Episodes", "Therapy", "Clinical", "Functional", "Service"),
    sep = ","
  )

hhrg2 <- new_pro_hhrg %>%
  mutate(
    Episodes = str_trim(Episodes),
    Therapy = str_trim(Therapy),
    Clinical = str_trim(Clinical),
    Functional = str_trim(Functional),
    Service = str_trim(Service)
  )
hhrg2
```

## 4.5
* These rows have an additional "," at the end of the context, so they are discarded.  
I don't think it makes sense to drop these rows. Because we can eliminate the comma in those rows and keep the original information.
```{r}
provider_hhrg %>%
  filter(row_number() == 122 | row_number() == 5580 | row_number() == 5590)
```

## 4.6
'Description' and 'Clinical, Functional, and Service Levels' are columns in case_mix_weight I plan to use for merging.  
There are 153 distinct HHRG groups using these columns.  

```{r}
case_mix_weights %>%
  count("Description", "Clinical, Functional, and Service Levels")
case_mix_weights <- rename(case_mix_weights,
  CFS = "Clinical, Functional, and Service Levels"
)
```

## 4.7

```{r}
new_casemix <- case_mix_weights %>%
  separate(
    Description,
    into = c("Episodes", "Therapy"),
    sep = ","
  )

new_casemix <- new_casemix %>%
  separate(
    CFS,
    into = c("Clinical", "Functional", "Service"),
    sep = "[D-Z]"
  )

casemix14 <- new_casemix %>%
  mutate(
    Episodes = str_trim(Episodes),
    Therapy = str_trim(Therapy),
    Clinical = str_trim(Clinical),
    Functional = str_trim(Functional),
    Service = str_trim(Service)
  )
casemix14
```

## 4.8

**Clinical**, **Functional** and **Service** columns have the same values in both datasets.   
**Episodes** and **Therapy** columns have similar values, but require further cleanup.  

We can see that there is a value overlapping in Episodes in hhrg2 (Early or Late Episode) which is not shown in cmw. And in Therapy in cmw, values are divided into 9 ranges, which is different from 3 ranges in hhrg2.

```{r}
hhrg2 %>%
  count(Episodes)
```

```{r}
casemix14 %>%
  count(Episodes)
```
```{r}
hhrg2 %>%
  count(Therapy)
```
```{r}
casemix14 %>%
  count(Therapy)
```

```{r}
hhrg2 %>%
  count(Clinical)
```
```{r}
casemix14 %>%
  count(Clinical)
```
```{r}
hhrg2 %>%
  count(Functional)
```
```{r}
casemix14 %>%
  count(Functional)
```
```{r}
hhrg2 %>%
  count(Service)
```
```{r}
casemix14 %>%
  count(Service)
```

I use fct_recode() to fix the columns in case_mix_weight to ensure that the five columns have identical values to provider_hhrg (hhrg2).
```{r}
cmw_clean <-
  casemix14 %>%
  mutate(Episodes = fct_recode(
    Episodes,
    "Early Episode" = "1st and 2nd Episodes",
    "Late Episode" = "3rd+ Episodes",
    "Early or Late Episode" = "All Episodes"
  ))
```

```{r}
cmw_clean <-
  cmw_clean %>%
  mutate(Therapy = fct_recode(
    Therapy,
    "0-13 therapies" = "0 to 5 Therapy Visits",
    "0-13 therapies" = "6 Therapy Visits",
    "0-13 therapies" = "7 to 9 Therapy Visits",
    "0-13 therapies" = "10 Therapy Visits",
    "0-13 therapies" = "11 to 13 Therapy Visits",
    "14-19 therapies" = "14 to 15 Therapy Visits",
    "14-19 therapies" = "16 to 17 Therapy Visits",
    "14-19 therapies" = "18 to 19 Therapy Visits",
    "20+ therapies" = "20+ Therapy Visits"
  ))
```
```{r}
hhrg_clean <-
  hhrg2 %>%
  mutate(Clinical = fct_recode(
    Clinical,
    C1 = "Clinical Severity Level 1",
    C2 = "Clinical Severity Level 2",
    C3 = "Clinical Severity Level 3"
  ))

hhrg_clean <-
  hhrg_clean %>%
  mutate(Functional = fct_recode(
    Functional,
    "1" = "Functional Severity Level 1",
    "2" = "Functional Severity Level 2",
    "3" = "Functional Severity Level 3"
  ))

hhrg_clean <-
  hhrg_clean %>%
  mutate(Service = fct_recode(
    Service,
    "1" = "Service Severity Level 1",
    "2" = "Service Severity Level 2",
    "3" = "Service Severity Level 3",
    "4" = "Service Severity Level 4",
    "5" = "Service Severity Level 5"
  ))
```


## 4.9
```{r}
cmw_clean <- cmw_clean %>%
  mutate(
    Episodes = as.character(Episodes),
    Therapy = as.character(Therapy)
  )

provider_hhrg_wt <- left_join(hhrg_clean, cmw_clean, by = c("Episodes", "Therapy", "Clinical", "Functional", "Service")) %>%
  mutate(Clinical = str_sub(Clinical, -1))
```

```{r}
test_that(
  "we have the right number of rows",
  expect_equal(nrow(provider_hhrg_wt), 111904)
)
```

```{r}
test_that(
  "we have no NA in casemix_2014",
  expect_true(sum(is.na(provider_hhrg_wt$casemix_2014)) == 0)
)
```


# 5 Billing Outlier Analysis
* Reference: https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/weighted.mean
* https://stackoverflow.com/questions/31944103/convert-currency-with-commas-into-numeric
```{r}
provider_hhrg_wt$Avg_Pymt_Amt_Per_Epsd <- parse_number(provider_hhrg_wt$Avg_Pymt_Amt_Per_Epsd)
provider_hhrg_wt$Tot_Epsd_Stay_Cnt <- parse_number(provider_hhrg_wt$Tot_Epsd_Stay_Cnt)
```

```{r}
provider_sum <- provider_hhrg_wt %>%
  select(Prvdr_ID, Prvdr_Name, State, Avg_Pymt_Amt_Per_Epsd, Tot_Epsd_Stay_Cnt, casemix_2014) %>%
  group_by(Prvdr_ID, Prvdr_Name, State) %>%
  summarise(
    mean_medi_pay = weighted.mean(Avg_Pymt_Amt_Per_Epsd, Tot_Epsd_Stay_Cnt),
    mean_casemix = weighted.mean(casemix_2014, Tot_Epsd_Stay_Cnt),
    total_epi = sum(Tot_Epsd_Stay_Cnt)
  )
```


## 5.1
* Question: How much variation is there in average cost per episode by home-health agency?


```{r}
mean(provider_sum$mean_medi_pay, na.rm = TRUE)
median(provider_sum$mean_medi_pay, na.rm = TRUE)
var(provider_sum$mean_medi_pay, na.rm = TRUE)
```

```{r}
ggplot(data = provider_sum) +
  geom_histogram(aes(mean_medi_pay),
    fill = "navy",
    color = "black",
    binwidth = 200,
    boundary = 0
  ) +
  scale_x_continuous(breaks = seq(0, 7000, by = 500)) +
  labs(
    x = "Average Cost per Episode",
    y = "Count",
    title = "Variation in average cost per episode by home-health agency"
  )
```
**Answer**
From the graph below, we can see the distribution of average cost per episode is right-skewed with the positive skewness. Combining the information of statistics, the variance of 541249 shows the large variation in values. The median of average cost per episode is smaller than the mean value. And values larger than 3250 in average cost per episode are more disperse.

## 5.2
* Question: How much variation is there in average cost after accounting for how sick patients are via casemix? Show three different ways to depict the covariation of these two variables. Then explain which plot you prefer to answer the question and why.


```{r}
cov(provider_sum$mean_casemix, provider_sum$mean_medi_pay)
cor.test(provider_sum$mean_casemix, provider_sum$mean_medi_pay)
```

```{r}
ggplot(provider_hhrg_wt) +
  geom_smooth(
    aes(x = casemix_2014, y = Avg_Pymt_Amt_Per_Epsd)
  ) +
  ggtitle(
    label = "Covariation in Average Cost after accounting for Patients' Sickness",
    subtitle = "Smooth"
  ) +
  labs(x = "Casemix in 2014", y = "Average Payment Amount per Episode")
```

```{r}
ggplot(provider_hhrg_wt) +
  geom_point(
    aes(x = casemix_2014, y = Avg_Pymt_Amt_Per_Epsd)
  ) +
  ggtitle(
    label = "Covariation in Average Cost after accounting for Patients' Sickness",
    subtitle = "Point"
  ) +
  labs(x = "Casemix in 2014", y = "Average Payment Amount per Episode")
```

```{r}
provider_hhrg_wt %>%
  mutate(casemix_2014 = cut_width(casemix_2014, 0.3)) %>%
  ggplot() +
  geom_boxplot(
    aes(x = casemix_2014, y = Avg_Pymt_Amt_Per_Epsd)
  ) +
  ggtitle(
    label = "Covariation in Average Cost after accounting for Patients' Sickness",
    subtitle = "Boxplot"
  ) +
  labs(x = "Casemix in 2014", y = "Average Payment Amount per Episode")
```
**Answer:**
I would use geom_smooth() as it is the best way to look at the covariation of the required variables. It uses a regression to answer the question. However, all the plots are informative in this case as we observe high correlation and covariance. 


## 5.3
* Question: Is there more dispersion in average cost or cost_normalized? Why does one have less dispersion than the other? Answer this question with a plot, not a statistic. You might find it helpful to find and link to a stack overflow thread on overlaying histograms with ggplot2?


```{r}
provider_sum <- provider_sum %>%
  mutate(
    cost_normalized = mean_medi_pay / mean_casemix
  )
provider_sum %>%
  ggplot() +
  geom_histogram(aes(x = cost_normalized),
    fill = "purple",
    alpha = 0.3, color = "black", position = "identity", binwidth = 100
  ) +
  geom_histogram(
    aes(x = mean_medi_pay),
    fill = "yellow",
    color = "black",
    alpha = 0.4,
    binwidth = 100
  ) +
  labs(title = "Variation in Normalized Cost (purple) & Average cost (yellow)",
       x = "Cost",
       y ="Count")
```
**Answer:**
https://stackoverflow.com/questions/6957549/overlaying-histograms-with-ggplot2-in-r was used for this answer. We see more variation in average cost than we do in normalized cost. It may be because that normalization makes outliers more concentrated to the center.

## 5.4
* Question: What are the top 5 HHAs with the highest billing per episode in Illinois? What might happen if OIG decided to try to push down costs at the 5 HHAs with the highest billing per episode in Illinois (unadjusted for case mix)? What are the top 5 HHAs with the highest markups (billing per episode after normalizing for case mix) in Illinois? Is there any overlap between these two lists?


```{r}
provider_sum %>%
  filter(State == "IL") %>%
  arrange(
    desc(mean_medi_pay)
  ) %>%
  head(5)

provider_sum %>%
  filter(State == "IL") %>%
  arrange(
    desc(cost_normalized)
  ) %>%
  head(5)
```
**Answer:**
The first list is produced as required along with the second list. If OIG decided to try to push down costs at the 5 HHAs with the highest billing per episode in Illinois (unadjusted for case mix), it would likely try to push down costs at facilities that most require that much payment per patient on average because of how expensive procedures are, which might be an indication as to how risky/vital they are. They would likely be making the wrong policy decision if they made this choice without accounting for normalized costs. We can see further proof of this as there is no overalp amongst the two lists of 5 providers each that we see.
