---
title: "applied_ps_3"
author: "Kevin Samsi, He Liu"
date: "09/05/2022"
output:
  pdf_document: default
  html_document:
    number_sections: yes
---
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(testthat)
library(fs)
library(rgdal)
library(tidycensus)
library(lubridate)
```

**Front matter**
This submission is my work alone and complies with the 30535 integrity policy.

Add your initials to indicate your agreement: **H.L.** & **K.S.**

Add names of anyone you discussed this problem set with: **Kevin Samsi**

Late coins used this pset: 0. Late coins left: 4. 

```{r}
setwd("/Users/8q48/Documents/GitHub/applied-ps-3-eva-kevin")
```


# 2 Read in one percent sample

## 2.1

It takes 1.503769 seconds to read in this file.

```{r}
start_time <- Sys.time()
tickets_1pct <- read_csv("parking_tickets_one_percent-3.csv")
end_time <- Sys.time()

runningtime <- end_time - start_time
runningtime

# Citation: https://www.r-bloggers.com/2017/05/5-ways-to-measure-running-time-of-r-code/
```

```{r}
test_that(
  "There are 287458 rows in parktickets file",
  expect_equal(nrow(tickets_1pct), 287458)
)
```

## 2.2

80 * 100 = 8000 MB  
Because the one percent sample is 80 MB, we would predict the size of the full data set is 8000 MB.

```{r}
file_size("parking_tickets_one_percent-3.csv")

# Citation: https://stackoverflow.com/questions/30580798/how-to-check-file-size-before-opening
```

## 2.3

The rows are ordered by the first column in the dataset, which is called "...1".

## 2.4

There are 97 NAs in `license_plate_state`, 2054 NAs in `license_plate_type`, 54115 NAs in `zipcode`, 29 NAs in `unit`, 84068 NAs in `notice_level` and 259899 NAs in `hearing_disposition`. There is no NA in other columns.

```{r}
colSums(is.na(tickets_1pct))

# Citation: https://stackoverflow.com/questions/26273663/r-how-to-total-the-number-of-na-in-each-col-of-data-frame
```

## 2.5
* There are more missing values in `notice_level` because the blank field means no notice was sent.  
* There are more missing values in `zipcode` because the ZIP code is associated with the vehicle registration. If the car is not registered, tben the value is NA for that car.  
* There are more missing values in `hearing_disposition` because if the ticket was not contested this field is blank which shows in NA.


# 3 Cleaning the data and benchmarking

## 3.1

* 22364 tickets were issued in 2017.  
22364 * 100 = 2,236,400  
It implies that 2,236,400 tickets were issued in 2017 in full data.  
* According to the ProPublica article, tickets issued annually were in ranges between 200,000 and 250,000 from 2007 to 2016.  
I think there is a meaningful difference because it is a ten times difference. It may because some dramatic changes happened during 2017 and the  intensity of punishment increased.
```{r}
tickets_1pct %>%
  separate(issue_date, c("date", "time"), sep = " ") %>%
  separate(date, c("year", "month", "day"), sep = "-") %>%
  filter(year == 2017) %>%
  count()
```

## 3.2

```{r}
tickets_1pct %>%
  group_by(violation_description) %>%
  summarise(n = n()) %>%
  arrange(desc(n)) %>%
  head(20) %>%
  mutate(violation_type = fct_reorder(violation_description, n)) %>%
  ggplot(aes(x = violation_type, y = n)) +
  geom_col(position = "identity") +
  theme(
    axis.text.y = element_text(size = 5.5)
  ) +
  coord_flip() +
  labs(
    x = "Frequency", y = "Violation type",
    title = "Top 20 most frequent violation types"
  )

# Citation: https://blog.51cto.com/matrix6ro/1793431
```


# 4 joins – unit

## 4.1

There are 29 missing values in unit.
```{r}
tickets_1pct %>%
  filter(is.na(unit))
```

## 4.2

There are 385 units in unit_key.csv either by canvas table or table for WBEZ.

```{r}
unit_key <- read_csv("unit_key-1.csv", skip = 2) %>%
  select(, -5)
names(unit_key) <- c(
  "Reporting District_canvas", "Department Name_canvas",
  "Department Description_canvas", "Department Category_canvas",
  "Reporting District_WBEZ", "Department Category_WBEZ"
)

length(unique(unit_key$`Reporting District_canvas`))
length(unique(unit_key$`Reporting District_canvas`))
```

## 4.3

* 129 rows in the tickets data have a match in the unit table.  
256 rows are unmatched.
* 287458 rows in the unit data have a match in the tickets table.  
0 rows are unmatched.
```{r}
tickets_1pct <- tickets_1pct %>% mutate(unit_chr = as.character(unit))
unit_match <- unit_key %>%
  semi_join(tickets_1pct, c("Reporting District_canvas" = "unit_chr"))
unit_unmatch <- unit_key %>%
  anti_join(tickets_1pct, c("Reporting District_canvas" = "unit_chr"))

t1pct_match <- tickets_1pct %>%
  semi_join(unit_key, c("unit_chr" = "Reporting District_canvas"))
t1pct_unmatch <- tickets_1pct %>%
  anti_join(unit_key, c("unit_chr" = "Reporting District_canvas"))
```

## 4.4
Department of Finance issues more tickets.  
Department of Finance issues 143909 tickets and Chicago Police issues 127079 tickets.
```{r}
tickets_1pct %>%
  count(unit_description == "DOF")
tickets_1pct %>%
  count(unit_description == "CPD" | unit_description == "CPD-Airport" |
    unit_description == "CPD-Other")
```


# 5 Joins – ZIP code

## 5.1

```{r}
CENSUS_KEY <- Sys.getenv("CENSUS_API_KEY")

chicago_zip <- get_acs(
  geography = "zcta",
  variables = c(
    population = "B01003_001",
    blackshare = "B02001_003",
    medincome = "B19013_001"
  ),
  state = "IL",
  year = 2014
)

chicago_zip <- chicago_zip %>%
  pivot_wider(names_from = "variable", values_from = c("estimate", "moe"))

```

## 5.2
```{r}
zipcode <- read_csv("chi_zips.csv") %>%
  mutate(zip = as.character(ZIP))

tickets_1pct_unit_key <- tickets_1pct %>%
  mutate(zipcode = substr(zipcode, 1, 5)) %>%
  semi_join(zipcode, c("zipcode" = "zip")) %>%
  left_join(chicago_zip, c("zipcode" = "GEOID"))
```

## 5.3

The three neighborhoods with the most unpaid tickets are Loop(60604), Englewood(60621) and Austin(60644).

```{r}
tickets_unpaid <- tickets_1pct_unit_key %>%
  group_by(zipcode) %>%
  mutate(paid = if_else(ticket_queue == "Paid", 0, 1)) %>%
  summarise(unpaid_rate = sum(paid)/mean(estimate_population)) %>%
  arrange(desc(unpaid_rate)) %>%
  mutate(zip_type = fct_reorder(zipcode, unpaid_rate)) %>%
  filter(!is.na(unpaid_rate))

tickets_unpaid %>% head(3)
```

## 5.4
```{r}
tickets_unpaid %>%
  filter(!is.na(unpaid_rate)) %>%
  left_join(chicago_zip, c("zipcode" = "GEOID"))
```

```{r}
map <- get_acs(
  geography = "zcta",
  variables = c(population = "B01003_001"),
  state = "IL",
  year = 2014,
  geometry = TRUE
) %>%
  right_join(tickets_unpaid, c("GEOID" = "zipcode"))
```
```{r,fig.height = 3}
map %>%
  ggplot(aes(fill = unpaid_rate)) +
  geom_sf() +
  labs(
    title = "Number of Unpaid Tickets per Resident by ZIP",
    fill = "Unpaid Rate of Tickets"
  ) +
  theme(plot.title = element_text(hjust = 0.5)) 
```

